Apache Storm

### 2018-01-22 paper.tex
Prepare for review with supervisor

\begin{figure}
  \begin{center}
    \includegraphics[width=.7\textwidth]{topo.png}
    \caption{Generic Storm Topology}
    \label{fig:topo}
   \end{center}
\end{figure}

@@

@@

@@

 \subsubsection{Development and Testing}
All three tools provide their own API for development and testing. A comparision among these tools can be made on a basis of the programming languages supported, Apache Samza currently only supports JVM languages, whereas Spark Streaming supports Java, Scala and Python. Apache Storm on the other hand supports JVM as well as Non JVM languages which increases development speed, as the developers can choose their language according to their comfort.

@@

 \subsubsection{Benchmarks}
There have not been a comparative benchmark between Apache Storm, Spark Streaming and Apache Samza, but various sources have different comparisions between tools.  \cite{benchmark} considers Apache Storm, Spark Streaming and Apache Flink. The benchmark design consisted of counting the number of events taking place in the stream. The design was setup to read data from a Kafka Event Queue, the data being processed and result updated into Redis. The result included the window start time and the window end time, thus giving  an idea about the total latency of processing. According to this benchmark, they deduced that Apache Storm and Flink had low latency, whereas Spark Streaming had higher maximum throughput depending upon the batch duration setting.

 \section{Discussion}

>>>>

@@

\begin{figure}
  \begin{center}
    \includegraphics[width=0.85\textwidth, height= 0.23\textheight]{topo.png}
    \caption{Generic Storm Topology}
    \label{fig:topo}
   \end{center}
\end{figure}

@@

\begin{figure}
  \begin{center}
    \includegraphics[width=\textwidth, height=0.25\textheight]{throughput.png}
    \caption{Throughput in time with respect to changing worker node availability \cite{stormtwitter}}
    \label{fig:throughput}
   \end{center}
\end{figure}

As we can see in figure \ref{fig:throughput}, an experiment conducted in \cite{stormtwitter} to check the resilience and performance of storm under loss of worker nodes. The worker nodes were turned off one by one in fixed durations, and the throughput of the system was measured at every instance. The durations and number of machines are shown below in table \ref{tab:experiment}. As seen, 3 machines were killed every 3 minutes, but the number of workers were limited to 50 at every instance, thus increasing the number of workers per machine. The graph \ref{fig:throughput} suggests that every 15 minutes, when the worker nodes were killed, the throughput was reduced significantly while storm tries to balance the topology again. The balancing does not require extensible amount of time and results into a new maximum throughput. As we can observe in the graph, the throughput decreses every time a node is removed, which is an expected phenomenon, due to the decrease in the computational capacity.

\begin{table}
\caption{Experimental setup during different time frames}
\label{tab:experiment}
\begin{center}
\begin{tabular}{r@{\quad}cl}
\hline
\multicolumn{1}{l}{\rule{0pt}{12pt} 
	Time}&\multicolumn{1}{l}{no of machines}\\[2pt]
\hline\rule{0pt}{12pt}
0 minutes  & 16  \\
+15 minutes & 13 \\
+30 minutes & 10 \\
+45 minutes & 7 \\
+60 mintues & 4 \\[2pt]
\hline
\end{tabular}
\end{center}
\end{table}

@@

 \subsubsection{Development and Testing}
All three tools provide their own API for development and testing. A comparision among these tools can be made on a basis of the programming languages supported, Apache Samza currently only supports JVM languages, whereas Spark Streaming supports Java, Scala and Python. Apache Storm on the other hand supports JVM as well as Non JVM languages which increases development speed, as the developers can choose their language according to their comfort. Apart from the fact that choosing a language for development, users can utilize existing code, thus allowing high portability.

@@

\begin{figure}
\begin{center}
\includegraphics[width=.75 \textwidth]{comparisiongraph.png}
\caption{Performance Comparision between Apache Storm, Apache Flink and Spark Streaming \cite{benchmark}}
\label{fig:comparisiongraph}
\end{center}
\end{figure}

 \subsubsection{Benchmarks}
There have not been a comparative benchmark between Apache Storm, Spark Streaming and Apache Samza, but various sources have different comparisions between tools.  \cite{benchmark} considers Apache Storm, Spark Streaming and Apache Flink. The benchmark design consisted of counting the number of events taking place in the stream. The design was setup to read data from a Kafka Event Queue, the data being processed and result updated into Redis. The result included the window start time and the window end time, thus giving  an idea about the total latency of processing. According to this benchmark, they deduced that Apache Storm and Flink had low latency, whereas Spark Streaming had higher maximum throughput depending upon the batch duration setting. Figure \ref{fig:comparisiongraph} shows the comparative analysis among Apache Flink, Spark Streaming and Apache Storm. The batch processing scheme of Spark Streaming allows it to process events with higher throughput but with a trade off that the processing is delayed with higher latency. In the figure, we can observe that storm has comparatively less delay in processing, but according to \cite{benchmark}, storm with acking beyond 135,000 events per second, could not keep up with the throughput.

 \section{Discussion}
Apache Storm acts as a real time stream processing tool which is capable of computing results with high reliability and minimum latency. The tool is comparatively easy to setup and configure, with out of the box configurations which is production ready. The storm API is relatively simple and does not have a complicated method calls, the programming is based on event driven programming and since storm supports read and writes based on standard IO, utilizing previously written code could be relatively easier than in the other available tools. Data is currently being generated and consumed at an unprecedented rate, thus stream processing technologies is a requirement and is gaining popularity and attention. With all the various tools available, one can decide on which tool best suites ones needs. According to the benchmarks, if minimum latency is required, Apache Storm or Apache Flink or any system with Continuous data flow model should be preferred. Whereas if high throughput is required with a non strict latency criteria, Spark streaming could prove to be a better option.
